---
title: "All Mocks are Wrong, but Some are Useful"
author: "Nicholas Vietto"
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
execute: 
  warning: false
---

## Full Description
This table presents the first-round results of the 2024 NFL Draft alongside the final 2024 NFL Mock Draft predictions from the four most popular draft analysts. Although mock drafts offer engaging content throughout the draft season, their accuracy typically declines after the top four picks.

## Goal
The purpose of this table was to experiment with {gt} and {gtExtra} to create an engaging data visual


## Information on Packages and Data
Data was collected and complied from [NFL.com](https://www.nfl.com/news/bucky-brooks-2024-nfl-mock-draft-3-0) and [Sharp Football](https://www.sharpfootballanalysis.com/analysis/mel-kiper-mock-draft-nfl/). 

This table was built in R using [{gt}](https://gt.rstudio.com/) and [{gtExtras}](https://jthomasmock.github.io/gtExtras/). 

Data wrangling was completed using [{janitor}](https://github.com/sfirke/janitor), [{readxl}](https://github.com/tidyverse/readxl), and [{tidyverse}](https://www.tidyverse.org/). 


## Load Packages and Data

```{r}

pacman::p_load(gt, gtExtras, janitor, tidyverse, readxl)

data <- read_excel("data.xlsx")


```

## Data Wrangling 

```{r}


data <- data |> 
  clean_names() |> 
  as_tibble() |> 
  arrange(actual) |> 
  mutate(across(5:9, as.numeric)) |> 
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) |> 
  pivot_longer(cols = -c(players, actual_1, who_was_correct, mean_projection),
               names_to = 'draft',
               values_to = 'position') |>
  group_by(players, actual_1, who_was_correct, mean_projection) |> 
  summarise(positions = list(position)) |> 
  ungroup() |> 
  mutate(mean_projection = as.numeric(mean_projection)) |> 
  mutate(mean_projection_1 = mean_projection) |> 
  mutate(change_score = abs(mean_projection_1 - actual_1)) |> 
  select(players, mean_projection, actual_1, change_score, positions, who_was_correct)


color_palette <- c("#3B9AB2", "#F21A00", "#FF0000")

```

## Table 

```{r}
data |> 
  arrange(actual_1) |> 
  gt() |> 
  tab_header(
    title = md('**All Mocks are Wrong, but Some are Useful**'),
    subtitle = md("*The final 2024 <span style='font-family: Times New Roman;'>**NFL**</span> Mock Drafts from Daniel Jeremiah (DJ), Peter Schrager (PS), Mel Kiper (MK), Bucky Brooks (BB)*")) |> 
  tab_options(
    heading.align = "Center",
    heading.title.font.size = px(24),
    heading.background.color = "#5BBCD6") |> 
  opt_table_font(font = google_font(name = "Playfair Display")) |> 
  cols_label(
    players = 'Player Name',
    actual_1 = 'Actual Draft Position',
    change_score = 'Difference',
    who_was_correct = 'Who Predicted Correctly',
    mean_projection = 'Average Projected Draft Position',
    positions = md('Projected Position <br> (DJ, PS, MK, BB, Actual)'), 
  ) |> 
  gt_plt_sparkline(column = positions, palette = c("#FF0000", "#00A08A", "#F2AD00", "#F98400", "#5BBCD6"))|> 
  gt_color_rows(
    columns = c(actual_1, mean_projection),
    domain = c(0, 100),
    palette = color_palette) |> 
  opt_stylize(style = 6) |> 
  gt_plt_bar_pct(change_score, fill="#5BBCD6", scaled = FALSE) |> 
  cols_align(align = "center", columns = c(mean_projection, actual_1, change_score)) |> 
  tab_source_note(source_note = ("Data: NFL.com and Sharp Football | Table created by Nicholas Vietto for the 2024 Posit Table Contest"))




```





